Sparsity Weak Lensing $3$-D Density Map Reconstruction: Reducing the line of sight smearing.

### 2020-11-27 doc/paper_ms_method_HSCY1/ms.tex
correct typo in the ms.tex

\begin{abstract}
A new method is developed to reconstruct $3$-D density contrast maps from photometric weak-lensing shear measurements.
The $3$-D density contrast maps are modeled as a summation of the NFW basis atoms which have $2$-D multi-scale NFW surface
density profiles on the transverse plane and $1$-D Dirac delta functions in the line of sight direction. With the prior
assumption that the density fields can be sparsely represented by the NFW basis atoms, the density fields are reconstructed 
using an oracle algorithm: adaptive lasso. Our method is tested with realistic simulations using the HSC-like shape estimation
error and photo-$z$ uncertainty.
Our findings are summarized as folllows: 
1) The solution of the standard lasso algorithm suffers from a smear of structure 
in the line of sight direction even in the absence of shape nosie, and the addaptive lasso algorithm significantly removes 
the line of sight smear of structure. 
2) The algorithm is able to detect halo with minimal mass limits of $10^{13.5} M_{\odot}/h$, $10^{14.3} M_{\odot}/h$, 
$10^{15.0} M_{\odot}/h$ for the low ($z<0.3$), median ($0.3\leq z< 0.6$) and high ($0.6\leq z< 0.9$) redshifts, respectively.
3) The estimated redshift of the halos detected from the reconstructed mass maps are
lower than the true redshfit by about $0.03$ for halos at low redshifts ($z\leq 0.4$). The relative redshfit bias is below 
$0.5\%$ for halos at $0.4<z\leq 0.85$.
\end{abstract}

@@

\begin{figure*}[!t]
    \includegraphics[width=1.\textwidth]{nfwlet-atom-2D.pdf}
    \caption{The smoothed pixelized basis atoms. The upper row shows the basis atoms in Fourier space
            and the lower row shows the basis atoms in Real space.
            The leftmost column is the point mass atom and the other columns are the multi-scale NFW atoms.
            The smoothing kernel is Gaussian with standard deviation of $1.5$ pixels.} \label{fig-atoms2D}
\end{figure*}

@@

Light from distant galaxies is distorted by the intervening inhomogeneous density distribution along the line of sight
due to the influence of gravity. As a result of the light distortion, the shapes of the background galaxies are coherently
distorted. Such effect, which is known as weak-lensing, imprints the information of foreground mass density distribution
to the background galaxy images and offers a direct probe into the mass density distribution in our universe
\citep[see][for recent reviews]{revKilbinger15,revRachel17}.

The expected shear measurements ($\gamma$) on distant galaxies are related to the
foreground density contrast field ($\delta$) through a linear transformation:
\begin{equation} \label{eq-intro-delta2shear}
 \gamma=\mathbf{T} \delta,
\end{equation}
where $\mathbf{T}$ is used to denote the linear transformation operator which includes not only physical lensing effect
but also systematic effects from observations (e.g. pixelization and smoothing of shear field in
transverse plane, photo-$z$ uncertainty).

Several large scale surveys target to study the weak-lensing effect at high precision level (e.g. HSC \citep{HSC1-data}, KIDS
\citep{KIDS13}, DES \citep{DES05}, LSST \citep{LSSTScienceBook}, Euclid \citep{Euclid2011}, WFIRST \citep{WFIRST15}).

The primary goal of most weak-lensing surveys is to constrain the cosmology model through $2$-point correlations. The studies
include galaxy$-$galaxy lensing which cross correlating the shear field ($\gamma$) with the positions of foreground galaxies
\citep{gglens-GAMA-Han2014,gglens-BossCFHTMore2015,gglens-DES1},
and cosmic shear which auto-correlates the shear measurements
\citep{cosmicShearRealKids450,cosmicShear-DES1,cosmicShear_HSC1_Chiaki2019,cosmicShear_HSC1_Hamana2019}.
Since shear is directly related to the matter distribution as shown in eq. (\ref{eq-intro-delta2shear}), Galaxy$-$galaxy lensing
probes into the correlation between the matter field and galaxy field, on the other hand, cosmic shear probes into the
auto-correlation of matter field.

The reconstruction of density map from shear measurements also receive considerable interest as it reaches the nonlinear scales.
$2$-D density map reconstruction which recover an integration of projected mass along the line of sight has been well studied
within the community \citep{massMap-KS1993,WL-massMap-Glimpse2D-Lanusse2016,sparseBaysianMassMap-Price2020}
and applied to large scale surveys \citep{HSC1-massMaps,massMapDES-Chang2018,DES-SV-massMap-sparsity}. However, the
reconstruction of $3$-D mass map is still a challenging task.

In order to fully reconstruct the $3$-D mass density distribution ($\delta$) from the photometric shear observations ($\gamma$),
the density contrast field is modeled as a summation of basis atoms in a model dictionary:
\begin{equation} \label{eq-intro-dict}
 \delta= \mathbf{\Phi} x,
\end{equation}
where $\mathbf{\Phi}$ is the transformation operator from the projection coefficient field to the density contrast,
and $x$ denotes the parameters. \citet{LSS-massMap-Wiener-Simon2009} reconstruct the density field
in Fourier space, which is equivalent to model the mass field with sinusoidial functions. On the other hand,
\citet{LSS-massMap-Glimpse3D-Leonard2014} models the mass field with Starlets \citep{Starlet-Starck2015}.

The projection coefficients are estimated through optimizing a regularized loss function. The estimator is
generally defined as
\begin{equation}
\hat{x}=\argmin_{x} \left\{ \frac{1}{2}\norm{\Sigma^{-\frac{1}{2}}(\gamma- \mathbf{T\Phi} x)}_2^2+ \lambda C(x) \right\},
\end{equation}
where $\norm{\Sigma^{-\frac{1}{2}}(\gamma - \mathbf{T\Phi} x)}_2^2$ is the $l^2$ chi-square term\footnote{weighted by
the inverse of the diagonal covariance matrix of error on the shear measurements ($\Sigma$).} measuring
the residuals between the prediction and the data, while $C(x)$ is the regularization term measuring the deviation of
the estimation of the parameter ($x$) from the prior assumptions. The $l^p$ norm is defined as
\begin{equation}
\norm{x}_p=(\sum_{i} |x|^p_i)^{\frac{1}{p}}.
\end{equation}
Such estimation prefers the parameters that are able to describe the observations and also align with the prior assumptions.
The regularization parameter $\lambda$ adjusts the relative weight between the observations and prior assumptions in the
optimization process.

\citet{LSS-massMap-Wiener-Simon2009} propose to use the Wiener filter, which is also known as $l^2$ ridge regulation
($C=\norm{x}^2_2$), to find a regularized solution in Fourier space. \citet{HSC1-massMaps} apply the method of
\citet{LSS-massMap-Wiener-Simon2009} to the first-year data of the Hyper Suprime-Cam Survey \citep{HSC1-data}.
However, the density maps reconstructed by this method suffer from a smearing along the line of sight direction with
a standard deviation of $\sigma_z=0.2 \sim 0.3$.

\citet{LSS-massMap-Glimpse3D-Leonard2014} propose to use a derivative version of $l^1$ lasso regulation ($C=\norm{x}^1_1$)
to find a sparse solution in the Starlets dictionary space \citep{Starlet-Starck2015}. \citet{LSS-massMap-Glimpse3D-Leonard2014}
apply a greedy coordinate descent algorithm, which selects the steepest coordinate in each iteration, to find the minima
of a non-convex loss function penalized with the firm thresholding function. \citet{LSS-massMap-Glimpse3D-Leonard2014}
significantly reduce the smearing along the line of sight, however, the stability of the non-convex optimization and
greedy coordinate descent algorithm has not been fully justified. Moreover, the Starlets functions are not designed to
model the profile of clumpy mass in the universe.

$N$-body simulations have shown that the dark matter is distributed in halos connected by filaments, and the density
profile of a single halo follows the NFW function \citep{halo-NFW1997ApJ}.
We construct a dictionary with the multi-scale NFW atoms.  The atoms follow multi-scale surface density profiles of the 
NFW functions \citep{haloModel-TJ2003-3pt} on the transverse plane. We follow \citet{LSS-massMap-Glimpse3D-Leonard2014}
to neglect the depth of halos since the resolution scale of the reconstruction in the line of sight direction is much
larger than the scales of halos. Therefore, we set the profile of the NFW atoms in the line of sight as the Dirac delta
function. The multi-scale NFW atoms are used to model the profiles of halos which is assumed to be sparsely distributed
in the universe. With the sparsity prior, the adaptive lasso regularization \citep{AdaLASSO-Zou2006} is used to reconstruct
the density field. We find that, in contrast to the standard lasso estimator which smears the structure along the line of sight,
the addaptive lasso is able to significantly reduce the smearing effect.

Comparing with \citet{LSS-massMap-Glimpse3D-Leonard2014}, our dictionary is built up to describe the clumpy mass
in the universe which has a clear physical motivation. Furthermore, the adaptive lasso algorithm is strictly convex and
can be directly optimized with the FISTA algorithm \citep{FISTA-Beck2009} without relying on any greedy coordinate
descent approaches. The stability of this convex optimization has been well studied.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.5\textwidth]{photo-z_binning.pdf}
 \caption{The source galaxies are binned into $10$ redshift bins according to their Machine Learning and photo-Z (MLZ) 
        best photo-$z$ estimation. The blue histogram is the normalized number distribution of the best photo-$z$ estimation. 
        The vertical dashed lines are the boundaries of the bins. The galaxies are equal-number binned.} \label{fig-bestpz}
\end{figure}

@@

Finally, we solve the mass reconstruction problem in section \ref{subsec:method-reconstruction} using the adaptive lasso
algorithm \citep{AdaLASSO-Zou2006} optimized with the FISTA alogrithm \citep{FISTA-Beck2009}.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.5\textwidth]{mlz-poz.pdf}
 \caption{The average PDF of MLZ photo-$z$ error for $10$ source redshift bins.}\label{fig-pdfpz}
\end{figure}

@@

Substitute comoving distance ($\chi$) with redshift ($z$), we have
\begin{equation}\label{eq-delta2kappa}
\kappa(\vec{\theta},z_s)=\int_0^{z_s} dz_l K(z_l,z_s)\delta(\vec{\theta},z_l).
\end{equation}
where $K(z_l,z_s)$ is the lensing kernel defined as
\begin{equation}
K(z_l,z_s) =
\begin{cases}
\frac{3H_0\Omega_M}{2 c} \frac{\chi_l \chi_{sl} (1+z_l)}{\chi_{s} E\left(z_l\right)} & (z_s>z_l),\\
0&(z_s \leq z_l),
\end{cases}
\end{equation}
where $E(z)$ is the Hubble parameter as a function of redshift, in units of $H_0$.

\begin{figure*}[!t]
 \centering
 \includegraphics[width=1.\textwidth]{lensing_kernel.pdf}
 \caption{The left panel shows the lensing kernels for five different lens redshifts. The dashed lines are the
        kernels for spectroscopic redshift which assumes that the redshifts of source galaxies are precisely
        estimated. The solid lines are for photometric redshifts which accounts for the influence of photometric
        redshift uncertainty. The other two panels show the correlation between lensing kernels of different
        lens redshifts. The middle panel is for spectroscopic redshift and the right panel is for photometric
        redshifts. The lensing kernels are normalized so that the diagonal elements of the correlation matrices
        equal one.}\label{fig-corlensKer}
\end{figure*}

@@

\begin{figure*}[!t]
\centering
\subfigure[lasso]{\includezraphics{delta-3-7-pz-nn-lasso.pdf}}
\subfigure[adaptive lasso]{\includezraphics{delta-3-7-pz-nn-alasso.pdf}}
\caption{The results of density map reconstructions with standard lasso (left) and adaptive lasso (right). The input density
        map is from a NFW halo with mass $M_{200}=10^{15} ~h^{-1}M_{\odot}$ at redshift $0.35$. The lasso reconstruction
        smear the density field along the line of sight direction (vertical direction in the plots). The vertical direction
        is the line of sight direction, and the lower boundaries and upper boundaries of the boxes are $0.01$ and $0.85$, 
        respectively. Noises on galaxy shape measurements are neglected in this simulation.} \label{fig-lassoVsadaLasso}
\end{figure*}

The density contrast field is modeled as a summation of basis atoms in the dictionary:
\begin{equation}\label{eq-x2delta}
\delta(\vec{r}) = \sum_{s=1}^{N} \int d^3 r' \phi_s(\vec{r}-\vec{r'}) x_s(\vec{r'}),
\end{equation}
where $\phi_s(\vec{r})$ are the basis atoms of the dictionary. The basis atoms have `$N$' different scale frames and the atoms
in each scale frame are shifted by $\vec{r'}$ to form basis at different positions. $x_s(\vec{r'})$ is the projection
coefficients of the density contrast field onto the basis atoms.

We propose to use the multi-scale NFW atoms, denoted as $\{\phi_1,...,\phi_N\}$, as the basis atoms of our dictionary .
On the transverse plane, the NFW atoms follow surface density profiles of the NFW halos
\citep{haloModel-TJ2003-3pt} with scale radius $\theta_\alpha$ and truncation radius $c \theta_\alpha$ ,
where $c$ is the concentration of the NFW halos.
As the scales of halos are much less than the reachable redshift resolution, we neglect the depth of halo on the
line of sight direction and set the profiles of the NFW atoms in the line of sight direction to $1$-D Dirac delta
functions as suggested by \citep{LSS-massMap-Glimpse3D-Leonard2014}.
The multi-scale NFW atoms are defined as
\begin{equation}
\begin{split}
\phi_\alpha(\vec{r}) =&\frac{f }{2 \pi \theta_\alpha^2 } F(|\vec{\theta}|/\theta_\alpha) \delta_D(z),\\
&  (s=1..N)
\end{split}
\end{equation}
where
\begin{equation}
F(x)=
\begin{cases}
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccosh \left(\frac{x^2+c}{x(1+c)}\right)}{(1-x^2)^{3/2}}  & (x<1),\\
\frac{\sqrt{c^2-1}}{3(1+c)} (1+\frac{1}{c+1}) & (x=1),\\
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccos\left(\frac{x^2+c}{x(1+c)}\right)}{(x^2-1)^{3/2}} & (1<x\leq c),\\
0& (x>c).
\end{cases}
\end{equation}
$f=1/[\ln (1+c)-c/(1+c)]$. In this work, we fix $c=4$ for the NFW atoms in different scale frames.

To simplify the notation, we define the projection coefficients as a column vector:
$x=\begin{pmatrix}
x_{0}\\
x_{1}\\
...\\
x_{N}
\end{pmatrix}$,
and the dictionary transform operator as a row vector:
\begin{equation}
\mathbf{\Phi}=\begin{pmatrix}
\int d^3r\phi_0(\vec{r}) ~\int d^3r \phi_1(\vec{r})~ ...~\int d^3r \phi_{N}(\vec{r})
\end{pmatrix}.
\end{equation}
We substitute eq. (\ref{eq-x2delta}) into eq.(\ref{eq-delta2gammat}) and get
\begin{equation}\label{eq-x2gammat}
\gamma_L=\mathbf{Q}\mathbf{\Phi} x.
\end{equation}

In this paper, dictionary constructed with point mass atoms is used to compare with the dictionary of multi-scale atoms.
The point mass atoms is a $3$-D Dirac function defined as follows
\begin{equation}
\phi_{\rm{PM}}(\vec{r})= \delta_D(\theta_1) \delta_D(\theta_2) \delta_D(z).
\end{equation}

The $2$-D profiles of the point mass atom and the multi-scale NFW atoms on the transverse plane are shown in Figure
\ref{fig-atoms2D}. The $1$-D slices of the profiles are demonstrated in Fiugre \ref{fig-atoms1D}. Note that these profiles
are smoothed with a Gaussian kernel and pixelized into evenly spaced grids. The smoothing operation is discussed in
Section \ref{subsec:method-smoothing} and the pixelization operation is discussed in Section \ref{subsec:method-pixel}.

@@

The photometric redshifts of source galaxies in the current large scale survey are estimated with a limited number of
board photometric bands (e.g. $9$ bands for KIDS$+$VIKING survey \citep{KIDS_VIKING-Hildebrant2020}, $5$ bands for DES
survey and HSC survey). As a result, the estimated redshifts of galaxies suffer from much larger uncertainties, compared
with redshifts estimated with spectroscopic observations. Such photo-$z$ uncertainty smears the lensing kernels statistically
since a galaxy with a best fit photo-$z$ estimation of $z_s$ has possibilities of being actually located at different
redshifts ($z$). The probability function for the photo-$z$ uncertainty is denoted as $P(z|z_s)$ and the expected shear
distortion on the galaxy is
\begin{equation}\label{eq-delta2gamma-poz}
\int dz_s P(z|z_s) \gamma_L(\vec{\theta},z_s).
\end{equation}
With the definition of photo-$z$ smearing operator:
\begin{equation}
\mathbf{P} = \int dz_s P(z|z_s),
\end{equation}
the photo-$z$ uncertainty changes the shear as
\begin{equation}
\gamma_L \rightarrow \mathbf{Q} \gamma_L,
\end{equation}
where we use $\rightarrow$ to denote the changes of shear by the systematic operator.

Figure \ref{fig-bestpz} shows the histogram of the best Machine Learning and photo-Z \citep[MLZ]{MLZ-TPZ2013} photometric redshift estimation 
from \cite{HSC1-photoz} for galaxies in the tract 9347 of the HSC S16A data release \citep{HSC1-data}. These Galaxies are divided 
into ten source galaxy bins according to the photo-$z$ best estimation and the boundaries of the bins are shown as vertical dash 
lines in Figure \ref{fig-bestpz}. Figure \ref{fig-pdfpz} shows the average probability density function (PDF) in for galaxies 
in each redshift bin.

>>>>

\begin{abstract}
A new method is developed to reconstruct $3$-D density contrast maps from
photometric weak-lensing shear measurements. The $3$-D density contrast maps
are modeled as a summation of the NFW basis atoms, which have $2$-D multi-scale
NFW surface density profiles on the transverse plane and $1$-D Dirac delta
functions in the line of sight direction. With the prior assumption that the
density fields have a sparse spatial distribution, the density fields are
reconstructed using an oracle algorithm: adaptive lasso. Our method is tested
with realistic simulations using the HSC-like shape estimation error and
photo-$z$ uncertainty.  Our findings are summarized as follows: 1) The lasso
solution suffers from a smear of structure in the line of sight direction even
in the absence of shape noise. The adaptive lasso algorithm significantly
removes the line of sight smear of structure.  2) The algorithm is able to
detect halo with minimal mass limits of $10^{13.5} M_{\odot}/h$, $10^{14.3}
M_{\odot}/h$, $10^{15.0} M_{\odot}/h$ for the low ($z<0.3$), median ($0.3\leq
z< 0.6$) and high ($0.6\leq z< 0.9$) redshifts, respectively.  3) The estimated
redshift of the halos detected from the reconstructed mass maps are lower than
the true redshift by about $0.03$ for halos at low redshifts ($z\leq 0.4$). The
relative redshift bias is below $0.5\%$ for halos at $0.4<z\leq 0.85$.
\end{abstract}

@@

\begin{figure*}[!t] \includegraphics[width=1.\textwidth]{nfwlet-atom-2D.pdf}
    \caption{The smoothed pixelized basis atoms. The upper row shows the basis
        atoms in Fourier space, and the lower row shows the basis atoms in Real
        space.  The leftmost column is the point mass atom, and the other
        columns are the multi-scale NFW atoms.  The smoothing kernel is
        Gaussian with a standard deviation of $1.5$ pixels.
        } \label{fig-atoms2D}
\end{figure*}

@@

Light from distant galaxies is distorted by the intervening inhomogeneous
density distribution along the line of sight due to the gravity's influence.
As a result of the light distortion, the shapes of the background galaxies are
coherently distorted. Such effect, which is known as weak-lensing, imprints the
information of foreground mass density distribution to the background galaxy
images and offers a direct probe into the mass density distribution in our
universe \citep[see][for recent reviews]{revKilbinger15,revRachel17}.

The expected shear measurements ($\gamma$) on distant galaxies are related to
the foreground density contrast field ($\delta$) through a linear
transformation: \begin{equation} \label{eq-intro-delta2shear} \gamma=\mathbf{T}
\delta, \end{equation} where $\mathbf{T}$ is used to denote the linear
transformation operator, which includes not only physical lensing effect but
also systematic effects from observations (e.g., pixelization and smoothing of
the shear field in the transverse plane, photo-$z$ uncertainty).

Several large-scale surveys target to study the weak-lensing effect at high
precision level (e.g., HSC \citep{HSC1-data}, KIDS \citep{KIDS13}, DES
\citep{DES05}, LSST \citep{LSSTScienceBook}, Euclid \citep{Euclid2011}, WFIRST
\citep{WFIRST15}).

The primary goal of most weak-lensing surveys is to constrain the cosmology
model through $2$-point correlations. The studies include galaxy$-$galaxy
lensing, which cross-correlating the shear field ($\gamma$) with the positions
of foreground galaxies
\citep{gglens-GAMA-Han2014,gglens-BossCFHTMore2015,gglens-DES1}, and cosmic
shear which auto-correlates the shear measurements
\citep{cosmicShearRealKids450,cosmicShear-DES1,cosmicShear_HSC1_Chiaki2019,cosmicShear_HSC1_Hamana2019}.
Since the shear is directly related to the foreground matter distribution as
shown in eq.  (\ref{eq-intro-delta2shear}), Galaxy$-$galaxy lensing probes into
the correlation between the matter field and galaxy field. On the other hand,
cosmic shear probes into the auto-correlation of matter field.

The reconstruction of density map from shear measurements also receive considerable interest as it reaches the nonlinear scales.
$2$-D density map reconstruction which recover an integration of projected mass along the line of sight has been well studied
within the community \citep{massMap-KS1993,WL-massMap-Glimpse2D-Lanusse2016,sparseBaysianMassMap-Price2020}
and applied to large-scale surveys \citep{HSC1-massMaps,massMapDES-Chang2018,DES-SV-massMap-sparsity}. However, the
reconstruction of $3$-D mass map is still a challenging task.

To fully reconstruct the $3$-D mass density distribution ($\delta$) from the
photometric shear observations ($\gamma$), the density contrast field is
modeled as a summation of basis atoms in a model dictionary:
\begin{equation} \label{eq-intro-dict}
 \delta= \mathbf{\Phi} x,
\end{equation}
where $\mathbf{\Phi}$ is the transformation operator from the projection
coefficient field to the density contrast, and $x$ denotes the parameters.
\citet{LSS-massMap-Wiener-Simon2009} reconstruct the density field in Fourier
space, which is equivalent to model the mass field with sinusoidal functions.
On the other hand, \citet{LSS-massMap-Glimpse3D-Leonard2014} models the mass
field with Starlets \citep{Starlet-Starck2015}.

The projection coefficients are estimated by optimizing a regularized loss
function. The estimator is generally defined as
\begin{equation}
\hat{x}=\argmin_{x} \left\{ \frac{1}{2}\norm{\Sigma^{-\frac{1}{2}}(\gamma- \mathbf{T\Phi} x)}_2^2+ \lambda C(x) \right\},
\end{equation}
where $\norm{\Sigma^{-\frac{1}{2}}(\gamma - \mathbf{T\Phi} x)}_2^2$ is the
$l^2$ chi-square term\footnote{weighted by the inverse of the diagonal
covariance matrix of the error on the shear measurements ($\Sigma$).} measuring
the residuals between the prediction and the data, while $C(x)$ is the
regularization term measuring the deviation of the estimation of the parameter
($x$) from the prior assumptions. The $l^p$ norm is defined as
\begin{equation}
\norm{x}_p=(\sum_{i} |x|^p_i)^{\frac{1}{p}}.
\end{equation}
Such estimation prefers the parameters that are able to describe the
observations and align with the prior assumptions.  The regularization
parameter $\lambda$ adjusts the relative weight between the observations and
prior assumptions in the optimization process.

\citet{LSS-massMap-Wiener-Simon2009} propose to use the Wiener filter, which is
also known as $l^2$ ridge regulation ($C=\norm{x}^2_2$), to find a regularized
solution in Fourier space. \citet{HSC1-massMaps} apply the method of
\citet{LSS-massMap-Wiener-Simon2009} to the first-year data of the Hyper
Suprime-Cam Survey \citep{HSC1-data}.  However, the density maps reconstructed
by this method suffer from smearing along the line of sight direction with a
standard deviation of $\sigma_z=0.2 \sim 0.3$.

\citet{LSS-massMap-Glimpse3D-Leonard2014} propose to use a derivative version
of $l^1$ lasso regulation ($C=\norm{x}^1_1$) to find a sparse solution in the
Starlets dictionary space \citep{Starlet-Starck2015}.
\citet{LSS-massMap-Glimpse3D-Leonard2014} apply a greedy coordinate descent
algorithm, which selects the steepest coordinate in each iteration, to find the
minima of a non-convex loss function penalized with the firm thresholding
function. \citet{LSS-massMap-Glimpse3D-Leonard2014} significantly reduce the
smearing along the line of sight. However, the stability of the non-convex
optimization and greedy coordinate descent algorithm has not been fully
justified. Moreover, the Starlets functions are not designed to model the
profile of clumpy mass in the universe.

$N$-body simulations have shown that the dark matter is distributed in halos
connected by filaments, and the density profile of a single halo follows the
NFW function \citep{halo-NFW1997ApJ}.  We construct a model dictionary with the
multi-scale NFW atoms.  The atoms follow multi-scale surface density profiles
of the NFW functions \citep{haloModel-TJ2003-3pt} on the transverse plane.
Following \citet{LSS-massMap-Glimpse3D-Leonard2014}, we neglect the depth of
halos since the resolution scale of the reconstruction in the line of sight
direction is much larger than the halo scale. Therefore, we set the NFW atoms'
profile in the line of sight direction as the Dirac delta function.  Moreover,
we assume that the halos are sparsely distributed in the universe.  With the
sparsity prior, the adaptive lasso regularization \citep{AdaLASSO-Zou2006} is
used to reconstruct the density field.  In contrast to the lasso
estimator that smears the structure along the line of sight, the adaptive lasso
can significantly reduce the smearing effect.

Compared with \citet{LSS-massMap-Glimpse3D-Leonard2014}, our dictionary is
built up to describe the clumpy mass in the universe with a clear physical
motivation. Furthermore, the adaptive lasso algorithm is strictly convex and
can be directly optimized with the FISTA algorithm \citep{FISTA-Beck2009}
without relying on any greedy coordinate descent approaches. The stability of
this convex optimization has been well studied.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.5\textwidth]{photo-z_binning.pdf}
 \caption{The source galaxies are binned into $10$ redshift bins according to
         their Machine Learning and photo-Z (MLZ) best photo-$z$ estimation. The
         blue histogram is the normalized number distribution of the best photo-$z$
         estimation. The vertical dashed lines are the boundaries of the redshift bins.
         The galaxies are equal-number binned.
        } \label{fig-bestpz}
\end{figure}

@@

Finally, we solve the mass reconstruction problem in section
\ref{subsec:method-reconstruction} using the adaptive lasso algorithm
\citep{AdaLASSO-Zou2006} optimized with the FISTA algorithm
\citep{FISTA-Beck2009}.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.5\textwidth]{mlz-poz.pdf}
 \caption{The average PDF of MLZ photo-$z$ error for $10$ source redshift bins.
        }\label{fig-pdfpz}
\end{figure}

@@

After substituting comoving distance ($\chi$) with redshift ($z$), we have
\begin{equation}\label{eq-delta2kappa}
\kappa(\vec{\theta},z_s)=\int_0^{z_s} dz_l K(z_l,z_s)\delta(\vec{\theta},z_l).
\end{equation}
where $K(z_l,z_s)$ is the lensing kernel defined as
\begin{equation}
K(z_l,z_s) =
\begin{cases}
\frac{3H_0\Omega_M}{2 c} \frac{\chi_l \chi_{sl} (1+z_l)}{\chi_{s} E\left(z_l\right)} & (z_s>z_l),\\
0&(z_s \leq z_l),
\end{cases}
\end{equation}
where $E(z)$ is the Hubble parameter as a function of redshift, in units of $H_0$.

\begin{figure*}[!t]
 \centering
 \includegraphics[width=1.\textwidth]{lensing_kernel.pdf}
 \caption{The left panel shows the lensing kernels for five different lens
         redshifts. The dashed lines are the kernels for spectroscopic redshift,
         which assumes that the source galaxies' redshifts are precisely
         estimated. The solid lines are for photometric redshifts, which
         accounts for the influence of photometric redshift uncertainty. The
         other two panels show the correlation between lensing kernels of
         different lens redshifts. The middle panel is for spectroscopic
         redshift, and the right panel is for photometric redshifts. The
         lensing kernels are normalized so that the diagonal elements of the
         correlation matrices equal one.
        }\label{fig-corlensKer}
\end{figure*}

@@

\begin{figure*}[!t]
\centering
\subfigure[lasso]{\includezraphics{delta-3-7-pz-nn-lasso.pdf}}
\subfigure[adaptive lasso]{\includezraphics{delta-3-7-pz-nn-alasso.pdf}}
\caption{The density map reconstructions with the lasso (left) and the adaptive
        lasso (right) algorithms. The input density map is from a NFW halo with
        mass $M_{200}=10^{15} ~h^{-1}M_{\odot}$ at redshift $0.35$. The lasso
        reconstruction smears the density field along the line of sight
        direction (vertical direction in the plots).  The vertical direction is
        the line of sight direction, and the lower boundaries and upper
        boundaries of the boxes are $z=0.01$ and $z=0.85$, respectively. Noises
        on galaxy shape measurements are neglected in this simulation.
        } \label{fig-lassoVsadaLasso}
\end{figure*}

The density contrast field is modeled as a summation of basis atoms in the
dictionary:
\begin{equation}\label{eq-x2delta}
\delta(\vec{r}) = \sum_{s=1}^{N} \int d^3 r' \phi_s(\vec{r}-\vec{r'}) x_s(\vec{r'}),
\end{equation}
where $\phi_s(\vec{r})$ are the basis atoms of the dictionary. The basis atoms
have `$N$' different scale frames, and the atoms in each scale frame are
shifted by $\vec{r'}$ to form models at different spatial positions.
$x_s(\vec{r'})$ is the projection coefficients of the density contrast field
onto the basis atoms.

We propose to use the multi-scale NFW atoms, denoted as
$\{\phi_1,...,\phi_N\}$, as the basis atoms of our dictionary.  On the
transverse plane, the NFW atoms follow surface density profiles of the NFW
halos \citep{haloModel-TJ2003-3pt} with scale radius $\theta_\alpha$ and
truncation radius $c \theta_\alpha$ , where $c$ is the concentration of the NFW
halos.  As the scales of halos are much smaller than the reachable redshift
resolution, we neglect the depth of halo on the line of sight direction and set
the NFW atoms' profiles in the line of sight direction to $1$-D Dirac delta
functions as suggested by \citep{LSS-massMap-Glimpse3D-Leonard2014}. The
multi-scale NFW atoms are defined as
\begin{equation}
\begin{split}
\phi_\alpha(\vec{r}) =&\frac{f }{2 \pi \theta_\alpha^2 }
F(|\vec{\theta}|/\theta_\alpha) \delta_D(z),\\
&  (s=1..N)
\end{split}
\end{equation}
where
\begin{equation}
F(x)=
\begin{cases}
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccosh
\left(\frac{x^2+c}{x(1+c)}\right)}{(1-x^2)^{3/2}}  & (x<1),\\
\frac{\sqrt{c^2-1}}{3(1+c)} (1+\frac{1}{c+1}) & (x=1),\\
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} +
\frac{\arccos\left(\frac{x^2+c}{x(1+c)}\right)}{(x^2-1)^{3/2}} & (1<x\leq c),\\
0& (x>c).
\end{cases}
\end{equation}
$f=1/[\ln (1+c)-c/(1+c)]$. In this work, we fix $c=4$ for the NFW atoms in
different scale frames.

To simplify the notation, we compress the projection coefficients into a column
vector:
$x=\begin{pmatrix}
x_{0}\\
x_{1}\\
...\\
x_{N}
\end{pmatrix}$,
and compress the dictionary transform operator to a row vector:
\begin{equation}
\mathbf{\Phi}=\begin{pmatrix}
\int d^3r\phi_0(\vec{r}) ~\int d^3r \phi_1(\vec{r})~ ...~\int d^3r \phi_{N}(\vec{r})
\end{pmatrix}.
\end{equation}
We substitute eq. (\ref{eq-x2delta}) into eq.(\ref{eq-delta2gammat}) and get
\begin{equation}\label{eq-x2gammat}
\gamma_L=\mathbf{Q}\mathbf{\Phi} x.
\end{equation}

In this paper, a dictionary constructed with point mass atoms is used to
compare with the dictionary of multi-scale atoms.  The point mass atoms is a
$3$-D Dirac function defined as follows
\begin{equation}
\phi_{\rm{PM}}(\vec{r})= \delta_D(\theta_1) \delta_D(\theta_2) \delta_D(z).
\end{equation}

The $2$-D profiles of the point mass atom and the multi-scale NFW atoms on the
transverse plane are shown in Figure \ref{fig-atoms2D}. The $1$-D slices of the
profiles are demonstrated in Figure \ref{fig-atoms1D}. Note that these profiles
are smoothed with a Gaussian kernel and pixelized into evenly spaced grids. The
smoothing operation is discussed in Section \ref{subsec:method-smoothing}, and
the pixelization operation is discussed in Section \ref{subsec:method-pixel}.

@@

The photometric redshifts of source galaxies in the current large-scale survey
are estimated with a limited number of board photometric bands (e.g., $9$ bands
for KIDS$+$VIKING survey \citep{KIDS_VIKING-Hildebrant2020}, $5$ bands for DES
survey and HSC survey). As a result, the estimated redshifts of galaxies suffer
from much larger uncertainties than redshifts estimated with spectroscopic
observations. Such photo-$z$ uncertainty smears the lensing kernels
statistically since a galaxy with a best fit photo-$z$ estimation of $z_s$ has
possibilities of being actually located at different redshifts ($z$).  The
probability function for the photo-$z$ uncertainty is denoted as $P(z|z_s)$,
and the expected shear distortion on the galaxy is
\begin{equation}\label{eq-delta2gamma-poz}
\int dz_s P(z|z_s) \gamma_L(\vec{\theta},z_s).
\end{equation}
With the definition of photo-$z$ smearing operator:
\begin{equation}
\mathbf{P} = \int dz_s P(z|z_s),
\end{equation}
the photo-$z$ uncertainty changes the shear as
\begin{equation}
\gamma_L \rightarrow \mathbf{Q} \gamma_L,
\end{equation}
where we use $\rightarrow$ to denote the changes of shear by the systematic
operator.

Figure \ref{fig-bestpz} shows the histogram of the best Machine Learning and
photo-Z \citep[MLZ]{MLZ-TPZ2013} photometric redshift estimation from
\cite{HSC1-photoz} for galaxies in the tract 9347 of the HSC S16A data release
\citep{HSC1-data}. These Galaxies are divided into ten source galaxy bins
according to the photo-$z$ best estimation, and the boundaries of the bins are
shown as vertical dashed lines in Figure \ref{fig-bestpz}. Figure \ref{fig-pdfpz}
shows the average probability density function (PDF) for galaxies in each
redshift bin.
